Graph Storage: How good is CSR really?

### 2017-12-10 Seminar Paper/paper.tex
Some spelling mistakes corrected

\begin{abstract}
  In the last decade, the data size is growing exponentially and processing those data is becoming a difficult problem. Nowadays researchers and industry have been interested in the analysis of the graph to get deep understanding of social networks. Many of these graphs used in industry have become very large, containing hundreds of millions of nodes and edges. In this paper, I propose Compressed Sparse Row(CSR) which is a fast graph container. My CSR implementation also supports updates on graph. On a static graph, CSR gives ~ up to 3x better performance over Adjacency List Implementation(AL) on Depth-First-Search(DFS) and Breadth-First-Search(BFS) and sometimes better performance on Dijkstra algorithm(Shortest path).
\end{abstract}
\textbf{Keywords:} Graph databases, Compressed Sparse Row

\section{Introduction}
Graph analysis have become more popular in the last few years. Graphs are usually used to represent social networks in which vertices indicate users and edges indicate friendship between users. Facebook has 1.39 billion  active users as of 12/2014 with more than 400 billion edges\cite{fb}. Therefore, processing big graphs is consuming more time as the number of users and relationships significantly increased. The proposed approach(CSR) gives slightly better performance over the other approach on DFS algorithm. BFS algorithm in general is faster than DFS as it doesn't have an additional overhead of recursion. CSR gives a bit much better on BFS. In the following subsections, I will give some details about CSR and two variants of AL. In the following, I will give some details about existing approaches in Section 2, then followed by implementation details(Section 3) and an evaluation(Section 4). In Section 5 I will give conclusion about how good CSR is really.

\section{Related Works}
STINGER\cite{stinger} is a mutable graph container which is capable of handling updates which can be inserted parallel. it uses a fixed-size chained buckets to keep edges. Due to its loosely synchronized parallel updates that only maintains physical consistency of the data structure, a query may see the graph in a state that never existed. Each bucket contains only edges of the same type.\\
LLAMA\cite{llama} is a new extention to CSR with the support of versioning. It provides consistent snapshots on the graph and allow multiple accesses to them. At first, updates are buffered in a change set and after some period of time creates a new snapshots with the updates on change set. The main drawback is that the method usually has many existing snapshots and accessing to the newer snapshots consume more time. it merges old snapshots and apply a delta to CSR. \\
ASGraph\cite{asgraph} is also a new extention to which reduce update time to very small latency by breaking up the edge list of CSR into multiple chunks and that support efficient mutation and employing an append-only scheme for updates.

@@

\section{Evaluation}
All three approches are tested on different datasets which are generated randomly. The number of users is 1,000,000. To make it as a real social network, the number of neighbors differs from 50 to 200. In the following subsections, I discuss in which scenario CSR gives better performance in terms of latency and memory consumption. My test machine is a single socket server computer equipped with Intel(R) Core(TM) i7-3930K CPU 6-core @ 3.20GHz and 64 GB of main memory. This server has 12 hardware cores (with hyperthreading enabled).

@@

\begin{center}
\begin{tabular}{ |p{2cm}||p{4cm}|p{4.5cm}|}
 \hline
\# neighrbors per node & mem(std:list)$\div$mem(CSR) & mem(std:vector)$\div$mem(CSR)\\
 \hline
 50   & 7.33&1.41\\
 65&   7.46  & 2.01  \\
 80 &7.55 & 1.66\\
 95    &7.62 & 1.42\\
 110&   7.67  & 1.23\\
 125& 7.7  & 1.15 \\
 140& 7.73  & 1.85\\
 155& 7.76  & 1.68\\
 170& 7.78  & 1.54\\
 185& 7.79  & 1.42\\
 200& 7.81  & 1.31\\
 \hline
\end{tabular}\\
\textbf{Table 1}: Ratio of memory consumption of CSR with other approaches.
\end{center}

>>>>

\begin{abstract}
  In the last decade, the data size is growing exponentially and processing those data is becoming a difficult problem. Nowadays researchers and industry have been interested in the analysis of the graph to get deep understanding of social networks. Many of these graphs used in industry have become very large, containing hundreds of millions of nodes and edges. In this paper, I compare Compressed Sparse Row(CSR) as a fast graph container to Adjacency List(AL) in terms of latency and memory consumption. My CSR implementation also supports updates on graph. On a static graph, CSR gives ~ up to 3x better performance over AL Implementation on Depth-First-Search(DFS) and Breadth-First-Search(BFS) and sometimes better performance on Dijkstra algorithm(Shortest path).
\end{abstract}
\textbf{Keywords:} Graph databases, Compressed Sparse Row

\section{Introduction}
Graph analysis have become more popular in the last few years. Graphs are usually used to represent social networks in which vertices indicate users and edges indicate friendship between users. Facebook has 1.39 billion  active users as of 12/2014 with more than 400 billion edges~\cite{fb}. Therefore, processing big graphs is consuming more time as the number of users and relationships significantly increased. The proposed approach(CSR) gives slightly better performance over the other approach on DFS algorithm. BFS algorithm in general is faster than DFS as it doesn't have an additional overhead of recursion. CSR gives a bit much better on BFS. In the following subsections, I will give some details about CSR and two variants of AL. In the following, I will give some details about existing approaches in Section 2, then followed by implementation details(Section 3) and an evaluation(Section 4). In Section 5 I will give conclusion about how good CSR is really.

\section{Related Works}
STINGER~\cite{stinger} is a mutable graph container which is capable of handling updates which can be inserted parallel. it uses a fixed-size chained buckets to keep edges. Due to its loosely synchronized parallel updates that only maintains physical consistency of the data structure, a query may see the graph in a state that never existed. Each bucket contains only edges of the same type.\\
LLAMA~\cite{llama} is a new extention to CSR with the support of versioning. It provides consistent snapshots on the graph and allow multiple accesses to them. At first, updates are buffered in a change set and after some period of time creates a new snapshots with the updates on change set. The main drawback is that the method usually has many existing snapshots and accessing to the newer snapshots consume more time. it merges old snapshots and apply a delta to CSR. \\
ASGraph~\cite{asgraph} is also a new extention to which reduce update time to very small latency by breaking up the edge list of CSR into multiple chunks and that support efficient mutation and employing an append-only scheme for updates.

@@

\section{Evaluation}
All three approches are tested on different datasets which are generated randomly. The number of users is 1,000,000. To make it as a real social network, the number of neighbors differs from 50 to 200. The neighbors are actually selected randomly. In the following subsections, I discuss in which scenario CSR gives better performance in terms of latency and memory consumption. My test machine is a single socket server computer equipped with Intel(R) Core(TM) i7-3930K CPU 6-core @ 3.20GHz and 64 GB of main memory. This server has 12 hardware cores (with hyperthreading enabled).

@@

\begin{center}
\begin{tabular}{ |p{2cm}||p{4cm}|p{4.5cm}|}
 \hline
\# neighbors per node & mem(std:list)$\div$mem(CSR) & mem(std:vector)$\div$mem(CSR)\\
 \hline
 50   & 7.33&1.41\\
 65&   7.46  & 2.01  \\
 80 &7.55 & 1.66\\
 95    &7.62 & 1.42\\
 110&   7.67  & 1.23\\
 125& 7.7  & 1.15 \\
 140& 7.73  & 1.85\\
 155& 7.76  & 1.68\\
 170& 7.78  & 1.54\\
 185& 7.79  & 1.42\\
 200& 7.81  & 1.31\\
 \hline
\end{tabular}\\
\textbf{Table 1}: Ratio of memory consumption of CSR with other approaches.
\end{center}
